A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-06 manuscript.Rmd
caution to not use t.test or wilcox test

At the same time require PACs more than "just" reproducibility.
In contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
We call this property "reusability".
This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].

@@

At the heart of a PAC lies the planned analysis.
We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
This analyitical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]); we merely choose it for didactic purposes because similar procedures a widely used [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].

@@

@@

>>>>

PACs come with a challange though, in contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
We call this property "reusability".
This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].

@@

At the heart of a PAC lies the planned analysis.
We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]

@@

This function is one possible translation of the verbal descriptions:

@@

One point of note about the analytical strategy and the function implementing it is that there are several translations into code that would fit the descriptions but would produce different results.
One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
Both are valid implementation of Mann–Whitney–Wilcoxon test but the first assumes equal variance while the second applies Welch's correction and hence is robust even with unequal variances [@zimmermanRankTransformationsPower1993].
Mentioning every such minute implementation detail is would result in overly verbose preregistrations but does change the outcome.

We also want to caution researchers to not use the here shown procedure.
We merely choose it for didactic purposes because it may be familiar to many researchers because the t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
This analytical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]);
